|

### 2024-08-20 reproducibility/manuscript/manuscript.qmd
docs(manuscript): add trajectory inference methods
Signed-off-by: Cameron Smith <cameron.ray.smith@gmail.com>

>>>>

## Trajectory inference {#sec-methods-trajectory-inference}

### Velocity vector field

The velocity-based vector fields were generated using the scVelo function call

```python
scvelo.tl.velocity_graph(adata)
scvelo.tl.velocity_embedding(
  adata, 
  vkey='velocity',
  basis='embed',
)
```

We used the default options for projecting the vector fields from scVelo models.
Unlike scVelo, Pyro-Velocity uses statistics derived from posterior samples of
the denoised spliced gene expression and posterior samples of the velocity
estimation for building the cell state transition matrix estimates using cosine
similarity. Pyro-Velocity uses the same projection method as scVelo for
projecting the transition matrix into the two-dimensional vector field on the
user-provided embedding space.

### Clonal progression vector field

No method exists for projecting clonal cell fates from the LARRY data and
associated subsets thereof into a vector field. Thus, we designed a strategy for
projecting clonal progression vector fields. Given cells sharing the same LARRY
barcode, we considered the vectors connecting the embedding centroid of cells at
consecutive time points and then leveraged Gaussian smoothing of nearest
centroid directions to represent the clonal progression vector field informed by
LARRY barcodes on a regular grid over the cell manifold.
